---
title: "Untitled"
author: "Rachel Byers"
date: "2026-01-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, results = "hide")
options(repos=c(CRAN="https://cran.r-project.org"))
```
**Lab Setup**
```{r}
library(tidyverse)
library(palmerpenguins)
library(viridis)
library(RColorBrewer)
library(colorspace)
library(ggrepel)
library(colorblindr)
library(remotes)
library(gghighlight)
library(ggplot2)
```
**1.1** Create a scatterplot showing bill length vs. bill depth, colored by species. Compare at least three different qualitative palettes:
```{r}
data(penguins)
summary(penguins)

LengthxBill1 <- ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm, color = species)) + geom_point(size = 3, alpha = 0.7) + labs(title = "Scatterplot of Bill Length vs. Bill Depth by Species", x = "Bill Length", y = "Bill Depth") + scale_color_manual(values = c("Adelie" = "orange", "Chinstrap" = "blue", "Gentoo" = "#CC79A7")) + theme_minimal()

LengthxBill1

LengthxBill2 <- ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm, color = species)) +
geom_point(size = 3, alpha = 0.7) +
labs(title = "Scatterplot of Bill Length vs. Bill Depth by Species", x = "Bill Length", y = "Bill Depth") +
scale_color_brewer(palette = "Set2") + theme_bw()
LengthxBill2

LengthxBill3 <- ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm, color = species)) +
geom_point(size = 3, alpha = 0.7) +
labs(title = "Scatterplot of Bill Length vs. Bill Depth by Species", x = "Bill Length", y = "Bill Depth") +
scale_color_viridis_d(option = "plasma") + theme_bw()
LengthxBill3

LengthxBill4 <- ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm, color = species)) +
geom_point(size = 3, alpha = 0.7) +
labs(title = "Scatterplot of Bill Length vs. Bill Depth by Species", x = "Bill Length", y = "Bill Depth") +
scale_color_OkabeIto() + theme_bw()
LengthxBill4

```
**Reflection**
Out of the four color pallettes I tried, I think the first, LengthXBill1, is the best for this scenario. The three colors I chose for this plot are recommended by the readings for the course to be compatible with all forms of color blindness.  

**1.2** Create a boxplot of body mass by island. Use a colorblind-safe palette from ColorBrewer.
```{r}

```





